% theory_standalone.tex --- CP2 compile harness for sections/theory.tex.
% Standalone wrapper that supplies the preamble the section assumes, so the section can be
% compiled/proofed in isolation. The real manuscript will \input{sections/theory} under its own
% preamble; this file exists only to verify the section compiles. Build: `tectonic theory_standalone.tex`.
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
% Overleaf font (standing convention) --- commented for local compile:
% \usepackage{ebgaramond}

\emergencystretch=2em  % repo convention (cf. Gauge) --- absorbs overfull boxes from long \texttt/math.

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\title{\textbf{Minos --- theory section (standalone proof)}}
\author{Avery Karlin}
\date{}

\begin{document}
\maketitle

\input{sections/theory}

\begin{thebibliography}{9}
\bibitem{lacostejulien2011} S.~Lacoste-Julien, F.~Husz\'ar, Z.~Ghahramani.
  Approximate inference for the loss-calibrated Bayesian. \emph{AISTATS}, 2011.
\bibitem{vadera2021} M.~P.~Vadera et al. Post-hoc loss-calibration for Bayesian neural networks.
  \emph{UAI}, 2021.
\bibitem{zhao2021} S.~Zhao, M.~Kim, R.~Sahoo, T.~Ma, S.~Ermon. Calibrating predictions to decisions:
  a novel approach to multi-class calibration (decision calibration). \emph{NeurIPS}, 2021.
\bibitem{vickers2006} A.~J.~Vickers, E.~B.~Elkin. Decision curve analysis: a novel method for
  evaluating prediction models. \emph{Medical Decision Making}, 26(6):565--574, 2006.
\bibitem{ovadia2019} Y.~Ovadia et al. Can you trust your model's uncertainty? Evaluating predictive
  uncertainty under dataset shift. \emph{NeurIPS}, 2019.
\bibitem{gauge} A.~Karlin. Distribution-free conformal coverage for IVIM parameter maps, and the
  identifiability wall in the pseudo-diffusion compartment (Gauge). Companion manuscript.
\bibitem{echo} A.~Karlin. Echo: labeled repeatability spot-checks for deployment validity.
  Companion work.
\end{thebibliography}

\end{document}
