% =====================================================================
% Supplementary Information — railing-first redraft (NMR in Biomedicine).
% Built with tectonic: `tectonic supplement.tex`
% All values trace to Gnomon (reproduction.json, conditional_coverage.json)
% and Sextant (railing_results.json); see ../NUMBERS_FROZEN.txt.
% =====================================================================
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{microtype}
\usepackage{setspace}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage[hidelinks]{hyperref}

\graphicspath{{../figures/manuscript/}}

\newcommand{\Dstar}{\ensuremath{D^{*}}}
\renewcommand{\arraystretch}{1.2}

\setstretch{1.15}
\setlength{\parskip}{0.5em}
\setlength{\parindent}{0pt}

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{document}

\begin{center}
{\Large\bfseries Supplementary Information\par}
\vspace{0.6em}
{\itshape Pseudo-diffusion is weakly identifiable in practice: a reproducible
boundary-railing signature of \Dstar{} in open in-vivo IVIM, and the shape-aware
intervals that resolve it\par}
\vspace{0.4em}
Avery Karlin, BA --- Columbia University (Applied Physics and Applied Mathematics) and
University of Colorado Boulder (Computer Science)
\end{center}

\vspace{0.5em}

All supplementary values trace to the seeded reproduction (Gnomon, seed 20260620;
reframe seed 20260621) and the cross-cohort railing analysis (Sextant, seed 20260613,
5000-resample bootstrap), regenerated by the one-command pipelines in the Data and code
availability statement.

% ----------------------------------------------------------------- S1
\section*{S1\quad Cross-cohort railing detail (Sextant)}

Table~S1 gives the full per-cohort railing breakdown underlying main-text Figure~1 and
Section~3.1: high-SNR voxel count, railing rate (tight bounds, with 95\% bootstrap CI),
rail direction (fraction railed to the lower vs upper bound), the generous-bound control
(``wide''), and the pre-registered replication verdict. Railing is predominantly to the
upper bound in the abdomen but to the lower bound under the sparse three-b liver scheme
that omits $b=0$, consistent with the loss of low-$b$ leverage on \Dstar{}. The
generous-bound control still rails a substantial fraction in every cohort where it
applies, ruling out a tight-bound artefact.

\begin{table}[htbp]
\centering
\caption{Per-cohort \Dstar{} railing (Sextant; $\texttt{rail\_tol}$ tight bounds; 95\%
bootstrap CI, 5000 resamples). ``Analyzed'' is the bootstrapped voxel count (liver
cohorts subsampled from millions of high-SNR voxels). Pre-registered replication
threshold: point $\ge 0.30$, CI lower $\ge 0.20$. The clean four-b-value liver cohort
(proper $b=0$) is the primary independent cross-organ test (REPLICATES-STRONG); the
sparser three-b-value cohort is supporting evidence only---it omits $b=0$, is normalised
by its lowest acquired $b$, and pools just three subjects---and is reported as REPLICATES
(no wide-bound control).}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{lrrcccl}
\toprule
Cohort (scheme) & High-SNR & Analyzed & Railed \% [95\% CI] & lower/upper & Wide \% & Verdict \\
\midrule
OSIPI abdomen, homogeneous ROI & 1{,}618 & 1{,}618 & 54.7 [52.2,\,57.1] & 16/39 & 32.1 & REPLICATES-STRONG \\
OSIPI abdomen, full ROI & 19{,}652 & 19{,}652 & 47.8 [47.1,\,48.5] & 5/43 & 27.4 & REPLICATES-STRONG \\
TCGA-LIHC liver, 4-b (0/50/500/800) & 648{,}265 & 40{,}000 & 43.7 [43.2,\,44.2] & 17/27 & 15.2 & REPLICATES-STRONG \\
TCGA-LIHC liver, 3-b (50/400/800) & 1{,}544{,}999 & 40{,}000 & 73.4 [73.0,\,73.9] & 73/1 & --- & REPLICATES \\
\bottomrule
\end{tabular}}
\par\vspace{0.3em}
{\footnotesize Per-subject three-b liver railing rates (0.66, 0.74, 0.75) are mutually
consistent. The Gnomon clean-room reference independently obtains 54.2\% [52.0,\,56.4] on
the full homogeneous ROI (n\,=\,1932) at $\texttt{rail\_tol}=10^{-3}$ (Table~S2), the
numeric source of truth for the lead claim.}
\end{table}

% ----------------------------------------------------------------- S2
\section*{S2\quad Railing sensitivity to tolerance and SNR selection (Gnomon)}

The lead railing number is reported with its selection sensitivity rather than pinned to
a single unstated cut (Table~S2). Loosening the rail tolerance from $10^{-3}$ to
$10^{-2}$ and restricting to the b0-SNR$>$25 subset both raise the rate, as expected, but
the phenomenon is present at every setting. For reference, on the synthetic nine-cell
substrate the NLLS \Dstar{} railing rate is 17.5\%; railing is markedly more frequent in
vivo, where SNR and model mismatch are worse than the controlled synthetic regime.

\begin{table}[htbp]
\centering
\caption{Gnomon \Dstar{} railing rate on the open OSIPI abdomen by rail tolerance and SNR
selection (95\% bootstrap CI where computed). Homogeneous ROI $=$ 1932 voxels;
b0-SNR$>$25 subset $=$ 799 voxels.}
\small
\begin{tabular}{llcc}
\toprule
Selection & Tolerance & Railed \% & 95\% CI \\
\midrule
Homogeneous ROI (n=1932) & $\texttt{rail\_tol}=10^{-3}$ (primary) & 54.2 & [52.0,\,56.4] \\
Homogeneous ROI (n=1932) & $\texttt{rail\_tol}=10^{-2}$ & 56.2 & [54.0,\,58.4] \\
b0-SNR$>$25 (n=799) & $\texttt{rail\_tol}=10^{-3}$ & 58.7 & --- \\
b0-SNR$>$25 (n=799) & $\texttt{rail\_tol}=10^{-2}$ & 61.2 & --- \\
\bottomrule
\end{tabular}
\end{table}

% ----------------------------------------------------------------- S3
\section*{S3\quad Full conditional \Dstar{} coverage, both SD conventions (Gnomon)}

Table~S3 is the complete per-true-\Dstar{}-tercile central-95\% \Dstar{} coverage table
behind main-text Figure~2 and Table~1, for the symmetric Laplace SD, symmetric MCMC SD,
and MCMC quantile intervals, under both the honest Cram\'er--Rao convention (recommended)
and the floored convention (shown only to explain the retired marginal headline). Two
reconstruction checks confirm the divergence is fully accounted for by the convention:
the honest-Laplace high-\Dstar{} tercile reproduces $\approx$0.63 (0.633) and the
floored-Laplace pooled value reproduces $\approx$0.68 (0.676), the latter matching a
previously reported marginal MCMC-SD figure of 0.67. The reproduction's automated
reconstruction check passes.

\begin{table}[htbp]
\centering
\caption{Per-true-\Dstar{}-tercile central-95\% \Dstar{} coverage (nominal 0.95), both SD
conventions, with 95\% bootstrap CIs. Tercile edges (true \Dstar{})
$0.0511,\,0.0689\times10^{-3}$ mm$^2$/s; 600 voxels per tercile, 1800 pooled.}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{llcccc}
\toprule
Estimator & Convention & low \Dstar{} & mid \Dstar{} & high \Dstar{} & pooled \\
\midrule
Laplace SD & honest & 0.91 [0.89,0.94] & 0.86 [0.83,0.89] & 0.63 [0.60,0.67] & 0.80 [0.78,0.82] \\
Laplace SD & floored & 0.87 [0.85,0.90] & 0.75 [0.72,0.78] & 0.41 [0.37,0.45] & 0.68 [0.65,0.70] \\
MCMC SD & honest & 0.95 [0.93,0.97] & 0.95 [0.94,0.97] & 0.81 [0.78,0.84] & 0.90 [0.89,0.92] \\
MCMC SD & floored & 0.92 [0.90,0.94] & 0.85 [0.83,0.88] & 0.65 [0.61,0.69] & 0.81 [0.79,0.83] \\
\rowcolor{yellow!25} MCMC quantile & honest (rec.) & 0.93 [0.90,0.95] & 0.97 [0.95,0.98] & 0.81 [0.78,0.84] & 0.90 [0.89,0.91] \\
\bottomrule
\end{tabular}}
\par\vspace{0.3em}
{\footnotesize Recommended reporting is the honest MCMC-quantile row. The marginal
(pooled) figures 0.80/0.90 are undramatic; the load-bearing failure is the conditional
high-\Dstar{} under-coverage (0.63 Laplace, 0.81 MCMC/quantile). The previously reported
marginal 0.30/0.67 do not reproduce and are retired (main text Section~3.2).}
\end{table}

% ----------------------------------------------------------------- S4
\section*{S4\quad Reproduction configuration and diagnostics}

The synthetic substrate is a nine-cell design (three ground-truth $(D,\Dstar{},f)$
triples $\times$ SNR $\in\{10,20,40\}$ $\times$ 200 Rician noise realisations), seed
20260620. The MCMC sampler (random-walk Metropolis; burn 1500, keep 2000, thin 2)
achieved per-truth acceptance rates 0.272, 0.280, 0.235, within the efficient range for
random-walk proposals. Coverage used nine quantile levels spanning 0.025 to 0.975
(central-95\% interval taken from the 0.025 and 0.975 quantiles). Bootstrap CIs used 2000
resamples (Gnomon reframe, seed 20260621) and 5000
resamples (Sextant, seed 20260613). The reproduction verdict is PARTIAL: the railing
diagnostic (T1) and the mechanism (K2 quantile, K3 flow) reproduce within their frozen
tolerances, while the two marginal coverage targets (T3a Laplace-SD, T3b MCMC-SD) diverge
for the documented SD-convention reason (Table~S3). The OSIPI archive is SHA-256-verified
(2a53054d\ldots b3e) on every run.

\end{document}
