% =====================================================================
% Cover letter VARIANT B (lead-with-replication) -- NMR in Biomedicine
% EIC: Professor John R. Griffiths
% Build: tectonic cover_letter_replication.tex
% =====================================================================
\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}

\signature{Avery Karlin, BA \\ Department of Applied Physics and Applied Mathematics,
Columbia University \\ \texttt{ak5232@columbia.edu}}
\address{Avery Karlin \\ Columbia University \\ New York, NY, USA}

\begin{document}
\begin{letter}{Professor John R.\ Griffiths \\ Editor-in-Chief, \textit{NMR in Biomedicine}}
\opening{Dear Professor Griffiths,}

I am pleased to submit ``Boundary-railing of conventional NLLS fits as an
assumption-free pseudo-diffusion identifiability diagnostic in IVIM MRI'' for
consideration in \textit{NMR in Biomedicine}.

What distinguishes this manuscript is the strength of its replication. The central
finding --- that a conventional box-constrained least-squares IVIM fit drives the
pseudo-diffusion coefficient $D^{*}$ onto a parameter bound in a large fraction of
voxels --- is established not once but across four cohorts and two independent
implementations. On the open OSIPI in-vivo abdomen scan the railed fraction is
54.7\% of high-SNR voxels (95\% CI 52.2--57.1); an independent clean-room
reimplementation, sharing no code with the original, reproduces a previously
reported figure (54.2\%, 95\% CI 52.0--56.4); the effect persists over the full
abdominal ROI (47.8\%); and it recurs on an entirely separate organ, scanner
manufacturer, and site --- TCGA-LIHC liver DWI --- at 43.7\% (clean 4-$b$ scheme)
and 73.4\% across three subjects (sparse 3-$b$ scheme), the sparser perfusion
sampling railing more, exactly as $D^{*}$ identifiability predicts. Pre-registered
replication thresholds, fixed before analysis, are cleared by every cohort.

This robustness is possible because the statistic assumes nothing: railing is read
directly off the optimiser's output and needs no ground truth and no noise model.
That is the deliberate re-aim of the work relative to an earlier, calibration-led
version. By promoting an assumption-free, cross-cohort-replicated phenomenon to the
primary claim, the manuscript answers the methods objection that a calibration
verdict is only as trustworthy as the simulator behind its reference truth --- the
railing result cannot be overextended in that way.

The calibration ruler is kept but scoped honestly: evaluated only on synthetic
ground truth, reported conditionally on the true $D^{*}$ regime and under a stated
Cram\'er--Rao convention (high-$D^{*}$ under-coverage of 0.63; a shape-aware
quantile interval and an amortized posterior as the resolution). The earlier
dramatic marginal headline is dropped and shown to be an artefact of an
undocumented SD convention, now fully documented along with all dataset
identifiers, fitting and training detail, and the Cram\'er--Rao approximation.

Every number traces to a single seeded reproduction that regenerates with one
command on openly licensed data (OSIPI TF2.4, Zenodo 14605039; TCGA-LIHC via The
Cancer Imaging Archive), so the replication can be re-run by any reader. The
manuscript fits both principal categories of \textit{NMR in Biomedicine} --- an
original in-vivo IVIM study and a methodological contribution --- and is
complementary to learned-uncertainty frameworks for IVIM such as Casali et al.\
(2026, this journal).

The manuscript is original, has not been published previously, and is not under
consideration elsewhere. The author declares no competing interests, takes full
responsibility for the integrity of the work, and includes a declaration of
AI-assisted writing. I would be glad to suggest reviewers and to answer any
questions.

\closing{Sincerely,}
\end{letter}
\end{document}
