\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ebgaramond}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{authblk}
\usepackage[colorlinks=true,linkcolor=teal,citecolor=teal,urlcolor=teal]{hyperref}

\input{numbers.tex}
\emergencystretch=2em   % avoid overfull boxes under ebgaramond justification

\renewcommand{\d}{D}                 % keep \d free of the babel accent
\newcommand{\Ds}{D^{*}}
\newcommand{\unit}[1]{\,\mathrm{#1}}

\title{\textbf{Boundary-railing of conventional NLLS fits as an assumption-free
pseudo-diffusion identifiability diagnostic in IVIM MRI}\\[3pt]
\large A bound-pinned $\Ds$ failure mode, re-centred from a calibration ruler and
replicated across open human-abdominal diffusion MRI}

\author[ ]{ }
\affil[ ]{ }
\date{}

\begin{document}
\maketitle

\begin{abstract}
\noindent
\textbf{Purpose.} Reported uncertainty in intravoxel incoherent motion (IVIM)
imaging has been critiqued as \emph{overextended}: any coverage or calibration
statement is only as trustworthy as the noise and forward model that generated
its reference truth. We answer that critique not by defending the calibration
argument but by promoting, as the primary contribution, a result that
\emph{cannot} be overextended because it assumes nothing. That $\Ds$ is weakly
identified is itself long known; what is new is not the phenomenon but an
assumption-free, optimizer-level statistic that neutralises this objection where
no calibration ruler can, together with its first systematic, pre-registered,
cross-cohort quantification and an independent clean-room reproduction.

\textbf{Methods.} In a conventional bounded non-linear least-squares (NLLS)
bi-exponential IVIM fit, the pseudo-diffusion coefficient $\Ds$ is so weakly
identified that the optimizer frequently terminates \emph{on} a parameter bound.
We read this boundary-railing directly off the optimizer's output --- no ground
truth, no noise model --- and quantify it with voxel-level bootstrap confidence
intervals on one open human-abdominal acquisition (OSIPI), its full ROI, and an
independent liver-DWI cohort (TCGA-LIHC). The calibration ruler (coverage,
expected calibration error, sharpness) is retained but demoted to a scoped
secondary diagnostic, evaluated only where ground truth exists, and under an
explicitly stated honest Cram\'er--Rao (CRLB) convention.

\textbf{Results.} On the curated homogeneous ROI, \HomoFrac\% of \HomoN{}
high-SNR voxels rail $\Ds$ (95\% CI [\HomoCIlo, \HomoCIhi]), reproducing a prior
report and corroborated by an independent clean-room reimplementation
(\gnomonRailFrac\% [\gnomonRailCIlo, \gnomonRailCIhi]). Railing generalises to the
full abdomen (\FullFrac\%), survives generous bounds and every SNR stratum, and
replicates on liver DWI (\LihcFourFrac\% at a clean 4-$b$ scheme;
\LihcThreeFrac\% across \LihcThreeSubj{} subjects at a sparse 3-$b$ scheme).
On synthetic ground truth the flag is actionable but scoped: a railed voxel's
$\Ds$ is unreliable with precision \flagPrecTwenty{} (against a \flagBaseTwenty\%
base rate) and excluding railed voxels lowers the retained map's pooled $\Ds$
error by \flagExclRedTwenty\%, yet the flag is specific rather than sensitive
(recall \flagRecTwenty), so it marks where $\Ds$ is untrustworthy without
detecting all such voxels.
As a scoped secondary finding, under the honest CRLB symmetric Gaussian intervals
under-cover $\Ds$ specifically in the high-$\Ds$ tercile (central-95\% coverage
\covLapHonHigh{} [\covLapHonHighLo, \covLapHonHighHi] for the Laplace SD), and an
amortized neural posterior is simultaneously better-calibrated and sharper than
the railed-NLLS baseline.

\textbf{Conclusion.} A large, reproducible fraction of conventional NLLS IVIM fits
rail to a parameter bound on open human-abdominal data. Because the statistic is
read off the optimizer without any calibration-trust assumption, it is immune to
the overextension objection; the calibration ruler remains useful but scoped.
\end{abstract}

\section{Introduction}

The clinically load-bearing question in quantitative diffusion MRI is not how
accurate a point estimate is, but whether a reported \emph{error bar} can be
believed. The intravoxel incoherent motion (IVIM) model~\cite{lebihan1988}
separates tissue diffusion from microcapillary perfusion through a
bi-exponential signal decay, and its perfusion-sensitive pseudo-diffusion
coefficient $\Ds$ is both the clinically interesting parameter and the
notoriously ill-conditioned one~\cite{koh2011,lemke2011}. A natural way to certify
an uncertainty estimate is a calibration ruler --- empirical coverage, expected
calibration error (ECE), and sharpness at fixed nominal levels --- that scores
whether a method's stated uncertainty is honest. A growing literature applies
exactly such rulers to classical Bayesian and learned IVIM
estimators~\cite{while2017,orton2014,barbieri2020,kaandorp2021,casali2026}.

The objection a sceptical reviewer raises against any such ruler is that it is
itself \emph{overextended}. Empirical coverage is computed against a reference
truth simulated under an assumed noise and forward model, so a coverage claim
inherits exactly the trust it was meant to certify: if the simulator is
misspecified, the calibration verdict is too. This is not a hypothetical concern
--- model misspecification is the central failure mode documented for
simulation-based inference~\cite{hermans2022,cannon2022,frazier2020}, and a
methods reviewer of an earlier version of this work flagged the calibration
contribution as resting on precisely this unverifiable assumption.

This paper takes the objection seriously and re-aims the contribution. We promote
to the primary claim a phenomenon that requires no calibration trust at all:
\textbf{boundary-railing} of the conventional NLLS IVIM fit. When the likelihood
is flat in $\Ds$, a box-constrained optimizer does not find an interior optimum
--- it slides to a constraint and stops. The fraction of voxels for which this
happens is a property of the estimator applied to the data; it is read directly
off the optimizer's output and carries no assumption about ground truth or noise.
It therefore cannot be overextended. We are explicit about what is new. That
$\Ds$ is weakly identified --- and that a bounded fit can consequently pin it to a
bound --- is long established~\cite{koh2011,lemke2011}; the contribution is not
that observation but an assumption-free, optimizer-level statistic that answers
the overextension objection no calibration ruler can, given here its first
systematic, pre-registered, cross-cohort quantification and an independent
clean-room reimplementation. We show that this railing fraction is large,
reproducible, generalises beyond a hand-curated ROI, survives deliberately
generous bounds and every SNR stratum, and replicates on an independent organ,
scanner, and site.

The calibration ruler is not discarded. It is kept as a clearly scoped
\emph{secondary} diagnostic, evaluated only on data with ground truth, and
reported \emph{conditionally} on the true $\Ds$ regime under a stated honest
CRLB convention --- a deliberately narrower claim than the marginal headline an
earlier draft carried. Within that scope it yields a real, reproducible finding:
symmetric Gaussian intervals under-cover $\Ds$ in the high-$\Ds$ regime, the same
weak-identifiability wall that drives railing. The contribution is thus a single
mechanism --- weak $\Ds$ identifiability --- observed two ways: as an
assumption-free fact about the optimizer (primary), and as a scoped calibration
consequence where ground truth permits it (secondary).

\section{Methods}
\label{sec:methods}

This section is complete from line one: it names every dataset with a stable
identifier, states every fitting and training detail, and documents both the
Cram\'er--Rao approximation and the railed-voxel standard-deviation (SD)
convention that earlier reviewers flagged as missing. A completeness checklist is
given in Table~\ref{tab:checklist}.

\subsection{Datasets}

\paragraph{Primary real cohort --- OSIPI abdomen.} The primary in-vivo cohort is
the open human-abdominal IVIM acquisition distributed with the OSIPI TF2.4 IVIM
code collection (Zenodo record \osipiRecord, DOI \osipiDOI, \osipiLicense;
Philips 3T)~\cite{osipidata2025}. The archive provides a
$144\times144\times21$ volume with \osipiNvol{} diffusion measurements over
\osipiNb{} unique $b$-values $\{0,10,20,30,40,50,75,100,150,250,400,600\}\unit{s/mm^2}$.
We analyse the archive's homogeneous-tissue mask (\osipiHomoN{} voxels; each
voxel $b_0$-normalised by its mean $b{=}0$ signal) and, separately, the full
abdominal ROI. The homogeneous ROI is the cohort on which a previously reported
\fashionReported\% NLLS $\Ds$-railing fraction was first computed.

\paragraph{Independent replication cohort --- TCGA-LIHC liver.} The independent
cohort is liver diffusion-weighted imaging from the TCGA-LIHC collection (TCIA;
DOI \tcgaDOI, \tcgaLicense; Siemens 1.5T)~\cite{tcgalihc2016} --- an unrelated
site, scanner manufacturer, and organ. We use \tcgaNseries{} series: a clean
4-$b$ scheme ($b{=}0,50,500,800\unit{s/mm^2}$; one subject) and a sparse clinical
3-$b$ scheme ($b{=}50,400,800\unit{s/mm^2}$, no $b{=}0$; \LihcThreeSubj{} subjects).

\paragraph{Synthetic substrate for the ruler.} The calibration ruler requires
ground truth and is therefore evaluated only on a synthetic digital reference
object. Ground-truth $(\d,\Ds,f)$ are drawn from documented physiological priors;
Rician noise is added at SNR $\in\{10,20,40\}$ with 200 noise realisations per
condition (a 9-cell design). The substrate is generated by an independent
read-only sibling module and is fully reproducible from its master seed.

\paragraph{Data and IP posture.} All imaging is open and download-on-demand
(OSIPI, \osipiLicense; TCGA-LIHC, \tcgaLicense), checksum-verified, and never
redistributed in-tree; only provenance manifests are committed. No proprietary or
clinical data are used.

\subsection{Forward model and estimators}

The bi-exponential IVIM model is
\begin{equation}
S(b)/S_0 = (1-f)\,e^{-bD} + f\,e^{-b(D+\Ds)},
\label{eq:ivim}
\end{equation}
fit per voxel over $(S_0,\d,f,\Ds)$ in scaled parameters so every fit variable is
$O(1)$, with an analytic Jacobian verified against finite differences.

\paragraph{NLLS and boundary-railing.} The point estimator is a box-constrained
four-parameter trust-region-reflective least-squares fit
(\texttt{scipy.optimize.least\_squares}, analytic Jacobian, \texttt{max\_nfev}
$=400$). The box follows the stated prior range. A parameter is \emph{railed} iff
$|\hat{x}-\text{bound}|/(\text{upper}-\text{lower}) < \varepsilon$; we report the
$\Ds$-railing rate at $\varepsilon=10^{-3}$ (primary) and $10^{-2}$ (sensitivity),
and under a deliberately widened box (``generous bounds''). The railing rate
needs no ground truth: it is a property of the estimator applied to the data.

\paragraph{Laplace and MCMC posteriors.} For the synthetic ruler only, two
classical posteriors are formed: a Laplace/asymptotic Gaussian at the MAP with the
CRLB covariance $\sigma^2(J^\top J)^{-1}$ (symmetric SD interval), and a
clean-room random-walk Metropolis sampler (per-voxel, vectorised; Gaussian
likelihood, known $\sigma$, uniform-box prior; burn 1500, keep 2000, thin 2),
from which we read both a symmetric SD interval and a 2.5/97.5 quantile interval.

\paragraph{Amortized neural posterior (NPE).} A conditional autoregressive
normalizing flow (5 affine layers, per-dimension MLP conditioners, standard-normal
base)~\cite{papamakarios2021,cranmer2020} is trained by neural posterior
estimation on 80\,000 $(\theta,x)$ pairs ($\theta\sim$ uniform-box,
SNR $\sim U(10,40)$; Adam, learning rate $10^{-3}$, 40 epochs, batch 512),
written from scratch with no external SBI library.

\subsection{The CRLB assumption and the railed-voxel SD convention}
\label{sec:crlb}

Two distinct modelling choices live here; both are stated, and the second is
decisive for the calibration headline, so it is documented in full.

\emph{(a) Gaussian/CRLB approximation.} The Laplace covariance
$\sigma^2(J^\top J)^{-1}$ is a Gaussian approximation to the posterior, weakest
precisely where $\Ds$ is most skewed --- low SNR and high $\Ds$, exactly where the
under-coverage sits. The MCMC results do not rely on it (the sampler targets the
exact posterior under the box prior).

\emph{(b) Railed-voxel SD convention.} A voxel whose $\Ds$ is railed carries no
local Fisher information about $\Ds$: the Jacobian column collapses and $J^\top J$
is singular. Two defensible conventions then exist and give \emph{opposite}
coverage. Under the \textbf{honest CRLB} (our default) the pseudo-inverse assigns
such a voxel a wide $\Ds$ interval --- the statistically correct admission that
$\Ds$ is undetermined there. Under an \textbf{overconfident floor} a railed voxel
is instead given a small floored $\Ds$ SD, so its interval is narrow and
under-covers ``by design.'' The convention is decisive: flooring drops pooled
Laplace $\Ds$ coverage from \covLapHonPool{} to \covLapFloPool, reconstructing a
dramatic marginal figure that is an artefact of the convention rather than a
property of the data. We report all calibration results under the honest CRLB and
show the floored convention only as a labelled illustration
(Section~\ref{sec:convention}).

\subsection{Calibration ruler and uncertainty on the numbers}

The ruler --- coverage, ECE (mean $|\text{empirical}-\text{nominal}|$ over a level
grid), sharpness (mean interval width) --- is re-derived from published
definitions, with a $\Ds$-tercile conditional-coverage probe added. Every
load-bearing number carries a bootstrap confidence interval (voxel-level
resampling for the real cohorts, \railNboot{} resamples, seed \railSeed;
noise-realisation resampling for the synthetic ruler). A directional gap is
reported as significant only if its bootstrap CI excludes zero. Pre-registered
replication thresholds (fixed before running): a cohort \emph{replicates} if the
railed fraction is $\ge\thrPoint\%$ \emph{and} its 95\% bootstrap lower bound is
$\ge\thrCIlo\%$.

\section{Results}

\subsection{Primary: boundary-railing across cohorts (assumption-free)}
\label{sec:primary}

On the curated homogeneous ROI, \HomoFrac\% of the \HomoN{} high-SNR voxels have a
bound-railed NLLS $\Ds$ (95\% CI [\HomoCIlo, \HomoCIhi]), matching the previously
reported \fashionReported\% on the same data (verdict \HomoVerdict). An
independent clean-room reimplementation --- a separate forward model, fitter, and
railing test sharing no code with the original --- obtains \gnomonRailFrac\%
([\gnomonRailCIlo, \gnomonRailCIhi]) on the same ROI, and \gnomonRailSnr\% on a
stricter $b_0$-SNR${>}25$ subset; the two CIs overlap, so the phenomenon is not an
implementation artefact.

Railing is not a property of a hand-chosen homogeneous patch. Extending to the
entire abdominal ROI (n${=}\FullN$), the rate is \FullFrac\% ([\FullCIlo,
\FullCIhi], verdict \FullVerdict). It also recurs on an independent organ,
scanner, and site: on the clean 4-$b$ liver scheme the railed fraction is
\LihcFourFrac\% ([\LihcFourCIlo, \LihcFourCIhi], n${=}\LihcFourN$, verdict
\LihcFourVerdict), and across \LihcThreeSubj{} subjects at the sparse 3-$b$
clinical scheme it is \LihcThreeFrac\% ([\LihcThreeCIlo, \LihcThreeCIhi],
n${=}\LihcThreeN$, verdict \LihcThreeVerdict). All four cohorts clear the
pre-registered replication thresholds (Table~\ref{tab:railing}).

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lrrrl}
\toprule
Cohort & $n$ (high-SNR) & Railed & 95\% CI & Verdict\\
\midrule
Abdomen, homogeneous ROI (OSIPI) & \HomoN & \HomoFrac\% & [\HomoCIlo, \HomoCIhi] & \HomoVerdict\\
Abdomen, full ROI (OSIPI) & \FullN & \FullFrac\% & [\FullCIlo, \FullCIhi] & \FullVerdict\\
Liver, 4-$b$ $0/50/500/800$ (TCGA-LIHC) & \LihcFourN & \LihcFourFrac\% & [\LihcFourCIlo, \LihcFourCIhi] & \LihcFourVerdict\\
Liver, 3-$b$ $50/400/800$ (TCGA-LIHC) & \LihcThreeN & \LihcThreeFrac\% & [\LihcThreeCIlo, \LihcThreeCIhi] & \LihcThreeVerdict\\
\bottomrule
\end{tabular}
\caption{Boundary-railing of the NLLS $\Ds$ across open human-abdominal cohorts
(seed \railSeed, \railNboot-resample bootstrap). Pre-registered thresholds:
railed $\ge\thrPoint\%$ with bootstrap lower bound $\ge\thrCIlo\%$. The
homogeneous-ROI rate matches the prior report of \fashionReported\% and is
independently corroborated by a clean-room reimplementation
(\gnomonRailFrac\%~[\gnomonRailCIlo,~\gnomonRailCIhi]).}
\label{tab:railing}
\end{table}

\subsection{Mechanism: why $\Ds$ rails}
\label{sec:mechanism}

Railing is structural weak identifiability of $\Ds$, not measurement noise. It
does not vanish with SNR --- it persists across every SNR stratum --- and it
survives deliberately generous bounds, falling but remaining a substantial
minority (\HomoWide\% on the homogeneous ROI, \FullWide\% on the full abdomen), so
it is not an artefact of tight box constraints. It is dominated by the
\emph{upper} $\Ds$ bound: the homogeneous ROI splits \HomoLower\% lower-rail /
\HomoUpper\% upper-rail, and the full ROI \FullLower\% / \FullUpper\%. This is the
high-$\Ds$ identifiability wall: the fast perfusion compartment is so weakly
constrained by the data that the optimizer is pushed to the ceiling. The
direction of the effect across acquisition schemes confirms the mechanism --- the
sparser the low-$b$ perfusion sampling, the worse $\Ds$ identifiability, and the
more the fit rails (the sparse 3-$b$ liver scheme rails most). Equation~\eqref{eq:ivim}
makes this concrete: $\Ds$ governs a fast decay visible only at low $b$, so once
the low-$b$ sampling is thin the likelihood is nearly flat in $\Ds$ and the
asymptotic (CRLB) variance --- which is exactly the curvature the optimizer
cannot find --- diverges.

\subsection{The railing flag is actionable: it flags voxels whose $\Ds$ is
untrustworthy (synthetic ground truth)}
\label{sec:flag}

The railing \emph{rate} is a population statistic; calling the flag a
\emph{diagnostic} earns the word only if railing, voxel by voxel, marks the fits
whose $\Ds$ cannot be recovered. That can be checked solely where ground truth
exists, so we test it on the same synthetic substrate used for the ruler
(\flagN{} truths per SNR drawn over the documented prior, fit with the identical
bounded NLLS and rail thresholds). A voxel's $\Ds$ is labelled \emph{unreliable}
by a deliberately truth-referenced, convention-free criterion --- normalised
recovery error $|\hat{\Ds}-\Ds|/\Ds>\flagTau$ (the point estimate off by more
than half) --- so no interval or SD convention enters.

Railing is a \emph{specific} flag. At SNR 20, representative of the
weak-identifiability regime the real abdominal data occupy, a railed voxel's
$\Ds$ is unreliable with precision \flagPrecTwenty{} [\flagPrecTwentyLo,
\flagPrecTwentyHi] against a base rate of \flagBaseTwenty\% --- a
\flagLiftTwenty$\times$ enrichment (Table~\ref{tab:flag}). Acting on the flag is
measurably useful: \emph{excluding} railed voxels lowers the pooled median
normalised $\Ds$ error of the retained map by \flagExclRedTwenty\% (from
\flagErrAllTwenty{} to \flagErrKeptTwenty), and the bootstrap CI on that
reduction excludes zero; the benefit recurs at SNR 10 and 40
(\flagExclRedTen\% and \flagExclRedForty\%, both CI-significant).

It is, deliberately stated, \emph{not} a sensitive detector. Its recall is only
\flagRecTwenty{} [\flagRecTwentyLo, \flagRecTwentyHi]: most unreliable voxels do
not rail, because $\Ds$ is weakly identified \emph{broadly} and a fit can carry a
badly recovered interior $\Ds$ without the optimizer reaching a bound. As a binary
classifier of unreliability the flag therefore sits only modestly above chance
(balanced accuracy \flagAuc), consistent with the near-chance area-under-curve of
railing scored against an above-median error split. The earned reading is
precise: a \emph{positive} rail is a high-precision, actionable flag that a voxel's
--- or, pooled, an acquisition's --- $\Ds$ is untrustworthy and should be excluded
or down-weighted; the \emph{absence} of railing is not a clean bill of health. We
claim ``diagnostic'' for \emph{flagging} untrustworthy $\Ds$, not for detecting
all of it, and not as a clinical-decision rule (Section~\ref{sec:discussion}).

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{5pt}
\begin{tabular}{rrrlll}
\toprule
SNR & Railed & Unreliable & Precision [95\% CI] & Recall [95\% CI] & $\Ds$-err red.\\
& (syn.) & base & {\footnotesize $P(\text{unrel.}\,|\,\text{railed})$} & {\footnotesize $P(\text{railed}\,|\,\text{unrel.})$} & (excl. railed)\\
\midrule
10 & \flagRailedTen\% & \flagBaseTen\% & \flagPrecTen{} [\flagPrecTenLo,\flagPrecTenHi] & \flagRecTen{} [\flagRecTenLo,\flagRecTenHi] & \flagExclRedTen\%\\
20 & \flagRailedTwenty\% & \flagBaseTwenty\% & \textbf{\flagPrecTwenty{} [\flagPrecTwentyLo,\flagPrecTwentyHi]} & \flagRecTwenty{} [\flagRecTwentyLo,\flagRecTwentyHi] & \flagExclRedTwenty\%\\
40 & \flagRailedForty\% & \flagBaseForty\% & \flagPrecForty{} [\flagPrecFortyLo,\flagPrecFortyHi] & \flagRecForty{} [\flagRecFortyLo,\flagRecFortyHi] & \flagExclRedForty\%\\
\bottomrule
\end{tabular}
\caption{Railing as a per-voxel flag of an unreliable $\Ds$ on synthetic ground
truth (\flagN{} truths per SNR; ``unreliable'' $=$ normalised $\Ds$ error
$>\flagTau$; seed \railSeed, bootstrap CIs). Railing is \emph{specific}
(precision well above base rate; railed voxels carry the larger error) but
\emph{not sensitive} (recall well below one): a positive flag is actionable ---
excluding railed voxels lowers the retained map's pooled $\Ds$ error by the last
column --- whereas a non-railed voxel is not certified reliable.}
\label{tab:flag}
\end{table}

\subsection{Secondary, scoped: the calibration ruler under the honest CRLB}
\label{sec:ruler}

The same weak-identifiability wall has a calibration consequence, which we report
only where ground truth exists (the synthetic substrate) and only conditionally on
the true $\Ds$ regime. This is the deliberately scoped secondary contribution; it
is undefined on the real abdominal scan, which has no ground-truth $(\d,\Ds,f)$,
and that limitation --- the very point the overextension critique raised --- is
the argument for leading with railing.

\paragraph{Symmetric Gaussian intervals under-cover $\Ds$ in the high-$\Ds$
regime.} Under the honest CRLB, central-95\% $\Ds$ coverage is near-nominal in the
low-$\Ds$ tercile but falls in the high-$\Ds$ tercile: \covLapHonHigh{}
[\covLapHonHighLo, \covLapHonHighHi] for the Laplace SD and \covMcHonHigh{}
[\covMcHonHighLo, \covMcHonHighHi] for the MCMC SD, versus \covLapHonLow{} and
\covMcHonLow{} respectively in the low-$\Ds$ tercile (Table~\ref{tab:coverage}).
The failure is identifiability-driven and conditional --- it is concentrated
exactly where railing concentrates --- not a uniform marginal collapse.

\paragraph{The interval shape, not a wider SD, restores marginal coverage.}
Reading the MCMC posterior's 2.5/97.5 quantile interval rather than a symmetric SD
interval restores near-nominal \emph{marginal} $\Ds$ coverage (\margMqDstar{}
[\margMqDstarLo, \margMqDstarHi]); $\d$ and $f$ are already near-nominal
(\margMqD, \margMqF). The right interval \emph{shape}, not a larger SD, is the fix.
Conditionally, however, a residual high-$\Ds$ gap remains even for the quantile
interval (\covMqHonHigh{} [\covMqHonHighLo, \covMqHonHighHi] in the high-$\Ds$
tercile): the residual is the identifiability wall itself, not an interval-shape
artefact.

\paragraph{An amortized posterior beats the railed NLLS on calibration and
sharpness.} On the synthetic held-out set the neural posterior is simultaneously
better-calibrated and sharper than the railed-NLLS baseline: $\Ds$ coverage
\flowCov{} versus \nllsCov{} (gap \gapCov{} [\gapCovLo, \gapCovHi]), ECE \flowEce{}
versus \nllsEce, and sharpness \flowSharp{} versus \nllsSharp. All three gaps'
bootstrap CIs exclude zero. The railed Gaussian's wide-but-miscentred intervals
lose on every axis: railing is not merely a point-estimate pathology but an
uncertainty pathology that a calibrated estimator avoids.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lcccc}
\toprule
Estimator (honest CRLB) & low $\Ds$ & mid $\Ds$ & \textbf{high $\Ds$} & pooled\\
\midrule
Laplace SD & \covLapHonLow{} [\covLapHonLowLo,\covLapHonLowHi] & \covLapHonMid{} [\covLapHonMidLo,\covLapHonMidHi] & \textbf{\covLapHonHigh{} [\covLapHonHighLo,\covLapHonHighHi]} & \covLapHonPool{} [\covLapHonPoolLo,\covLapHonPoolHi]\\
MCMC SD & \covMcHonLow{} [\covMcHonLowLo,\covMcHonLowHi] & \covMcHonMid{} [\covMcHonMidLo,\covMcHonMidHi] & \textbf{\covMcHonHigh{} [\covMcHonHighLo,\covMcHonHighHi]} & \covMcHonPool{} [\covMcHonPoolLo,\covMcHonPoolHi]\\
MCMC quantile & \covMqHonLow{} [\covMqHonLowLo,\covMqHonLowHi] & \covMqHonMid{} [\covMqHonMidLo,\covMqHonMidHi] & \covMqHonHigh{} [\covMqHonHighLo,\covMqHonHighHi] & \covMqHonPool{} [\covMqHonPoolLo,\covMqHonPoolHi]\\
\bottomrule
\end{tabular}
\caption{Per-true-$\Ds$-tercile central-95\% $\Ds$ coverage on the synthetic
substrate, under the honest CRLB convention (Section~\ref{sec:crlb}), with
bootstrap CIs. The under-coverage is conditional --- concentrated in the
high-$\Ds$ tercile --- and survives the quantile-shape fix.}
\label{tab:coverage}
\end{table}

\subsection{The convention reconstructs --- but does not justify --- a dramatic
marginal figure}
\label{sec:convention}

For completeness we document how a much more dramatic marginal under-coverage can
be produced. Under the overconfident floored convention
(Section~\ref{sec:crlb}b), pooled Laplace $\Ds$ coverage drops to \covLapFloPool{}
[\covLapFloPoolLo, \covLapFloPoolHi] and the high-$\Ds$ tercile to \covLapFloHigh{}
[\covLapFloHighLo, \covLapFloHighHi]. Flooring an unidentified voxel's SD reports
$\Ds$ as if it were precisely known --- it manufactures severity. We therefore do
\emph{not} report a marginal headline coverage; the honest, reproducible statement
is the conditional table above, and the floored figures appear only here, with the
convention named, to explain how a larger number can arise.

\section{Discussion}
\label{sec:discussion}

\paragraph{What is and is not claimed.} The primary claim is narrow and
assumption-free: a large, reproducible fraction of conventional NLLS IVIM fits
rail $\Ds$ to a bound on open human-abdominal data, across ROI, bounds, SNR, and
an independent liver cohort. We further show, where ground truth permits it, that
the flag is \emph{actionable}: railing marks voxels whose $\Ds$ is untrustworthy
with high precision (\flagPrecTwenty{} at SNR 20 versus a \flagBaseTwenty\% base
rate), and excluding railed voxels measurably improves the retained map's pooled
$\Ds$ error (Section~\ref{sec:flag}). That earns the word ``diagnostic'' in a
deliberately bounded sense: the flag is \emph{specific, not sensitive} (recall
\flagRecTwenty), so a positive rail flags untrustworthy $\Ds$ but the absence of a
rail does not certify a reliable one --- we claim a flag for untrustworthy $\Ds$,
not a complete detector of it. Two scope boundaries follow and are stated
explicitly. First, this railing flag is read off the conventional optimizer's
output and is \emph{distinct from} an out-of-distribution self-consistency gate
(a learned check on a posterior's own coherence); the two must not be conflated,
and we evaluate only the former. Second, the flag's \emph{clinical} utility ---
whether excluding or down-weighting railed voxels changes a downstream diagnosis
or patient outcome --- requires outcome evidence that no simulation can supply and
is therefore left to future work; we claim measurement reliability, not clinical
decision-making. The secondary claim is scoped to ground-truth data: under a
stated honest CRLB, symmetric Gaussian intervals under-cover $\Ds$ in the
high-$\Ds$ regime, a quantile interval fixes the marginal but not the conditional
gap, and an amortized posterior dominates the railed NLLS. We explicitly do
\emph{not} claim a dramatic marginal coverage collapse: such figures are
reconstructible only under an undocumented floored SD convention and are reported
here solely as a labelled illustration. We also make no claim about the
out-of-distribution self-consistency gate above, estimator timing, or in-vivo
brain held-out-$b$ recovery; those questions are outside the scope of the evidence
presented and are left to future work.

\paragraph{Differentiation from known $\Ds$ instability.} That $\Ds$ is unstable
and hard to fit is textbook~\cite{koh2011,lemke2011}; the contribution here is not
that observation but an assumption-free, optimizer-level railing statistic ---
immune to the overextension critique that calibration claims invite --- delivered
with its first systematic, pre-registered, cross-cohort quantification and an
independent clean-room reproduction.

\paragraph{Differentiation from learned uncertainty.} Casali et
al.~\cite{casali2026} characterise one supervised network's aleatoric/epistemic
uncertainty budget, validated on preclinical mouse brain; this work instead
documents an estimator-agnostic, real-data identifiability signature in
human-abdominal DWI, their own residual $\Ds$ overconfidence being the learned
echo of the same wall that rails the conventional fit.

\paragraph{Limitations and scope.} The open primary cohort is a single subject;
we make no population claim from it, reporting a within-subject voxel statistic
with a voxel-bootstrap CI and supporting generality by cross-cohort agreement
(homogeneous ROI, full ROI, and an independent liver cohort). The 3-$b$ liver
scheme lacks $b{=}0$ and is normalised by its lowest $b$, an approximation flagged
as such; the 4-$b$ liver cohort with $b{=}0$ is the clean independent test.
Railing magnitude depends on the bound width and the $b$-scheme density; we report
both tight and generous bounds and per-scheme results rather than a single number.
The calibration ruler is evaluated only on synthetic ground truth and is, by
construction, undefined on the real scans --- which is exactly why it is the
secondary, not the primary, contribution.

\section{Conclusion}

A large, reproducible fraction of conventional NLLS IVIM fits rail to a parameter
bound on open human-abdominal data, and the effect generalises across ROI, bounds,
SNR, and an independent liver cohort. Because it is read off the optimizer without
any calibration-trust assumption, boundary-railing is a primary claim immune to
the overextension objection that calibration arguments invite. The calibration
ruler remains useful but scoped: under an honest CRLB convention it shows a real,
conditional high-$\Ds$ under-coverage and an amortized-posterior advantage, with
every number traced to a seeded, openly reproducible computation.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{8pt}
\begin{tabular}{ll}
\toprule
Completeness item & Where addressed\\
\midrule
Dataset identifiers & \S\ref{sec:methods}: OSIPI Zenodo \osipiRecord; TCGA-LIHC \tcgaDOI\\
Acquisition / $b$-schemes & \S\ref{sec:methods}: \osipiNb{} unique $b$; liver 4-$b$ and 3-$b$ schemes\\
Full fitting (NLLS, MCMC) & \S\ref{sec:methods}: box/init/solver/railing; sampler spec\\
Full training (NPE) & \S\ref{sec:methods}: architecture, 80k sims, prior, epochs\\
CRLB approximation & \S\ref{sec:crlb}(a)\\
Railed-voxel SD convention & \S\ref{sec:crlb}(b); illustration \S\ref{sec:convention}\\
Calibration ruler definitions & \S\ref{sec:methods}\\
Uncertainty / CIs & \S\ref{sec:methods}: bootstrap, seed \railSeed\\
Claims scope (no overclaim) & \S\ref{sec:primary}--\ref{sec:convention}; Discussion\\
\bottomrule
\end{tabular}
\caption{Completeness checklist --- the methods items an earlier review flagged as
missing, each mapped to where it is now addressed.}
\label{tab:checklist}
\end{table}

\section*{Data availability and reproducibility}
All imaging is open and download-on-demand: the OSIPI abdomen acquisition (Zenodo
\osipiRecord, \osipiLicense) and the TCGA-LIHC liver series (TCIA, \tcgaDOI,
\tcgaLicense), each checksum-verified and never redistributed in-tree. Every
number in this manuscript is regenerated from seeded result files by a single
consistency gate that fails the build on any drift; the synthetic substrate, the
railing analysis, the classical and amortized posteriors, and the calibration
ruler are each independent, seeded reimplementations.

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
