\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage[T1]{fontenc}
\usepackage{newpxtext}
\usepackage{parskip}
\pagestyle{empty}
\newcommand{\Dstar}{D^{*}}
\setlength{\emergencystretch}{2em}

\begin{document}
\noindent June 16, 2026

\medskip
\noindent Dear Prof.\ Jezzard,

\medskip
Please consider the enclosed Research Article, ``Distribution-Free Conformal
Coverage for IVIM Parameter Maps, and the Identifiability Wall in the
Pseudo-Diffusion Compartment,'' for publication in \emph{Magnetic Resonance in
Medicine}.

The paper's central result is an information-theoretic one about IVIM itself:
high-pseudo-diffusion ($\Dstar$) regime-conditional coverage is bounded by an
identifiability limit. Across the label-free conditioning methods we tested, the
high-$\Dstar$ compartment under-covers; the barrier is conditioning on the true
$\Dstar$ regime --- a latent axis the data do not reveal --- the IVIM instance of
the known impossibility of distribution-free conditional coverage (Barber, Cand\`es,
Ramdas \& Tibshirani, 2021). A Fisher-information / Cram\'er--Rao analysis shows
$\Dstar$ point-identifiability collapses in the same regime (the bound reaches the
regime-bin width): a concomitant diagnostic, not the proven cause. No label-free
method we tested closes the gap --- including a deliberately biased, prior-shrinkage
estimator, which buys $\Dstar$ point precision but not conditional coverage --- and
the gap is not an artifact of our estimator. We characterize the limit rather than
overclaim a solution.

We reach that result using distribution-free conformal prediction as both instrument
and lens. Model-based IVIM uncertainty quantification --- probabilistic networks,
deep ensembles, mixture-density models, Bayesian fitting --- carries no
finite-sample coverage guarantee; this is, to our knowledge, the first work to bring
conformal prediction and conformalized quantile regression to the ill-posed IVIM
inverse problem, with validated finite-sample marginal coverage for $D$, $\Dstar$,
and $f$, and the first head-to-head benchmark against four model-based baselines.
Two practical findings follow: model-based IVIM UQ is overconfident broadly (not
specifically in $\Dstar$), and conformalizing a strong model-based base is the
sharpest valid recipe.

On scope, stated plainly: a coverage guarantee requires ground-truth parameters,
which in-vivo IVIM cannot supply, so realized coverage is unmeasurable in vivo in
principle. The guarantee is therefore established on a controlled synthetic cohort.
To address the natural concern that such a guarantee might be an artifact of our own
forward model, we replicate the marginal coverage and the high-$\Dstar$
identifiability wall on an external, community-standard digital reference object
(the OSIPI TF2.4 IVIM reference), on ground-truth parameters we did not generate ---
and, more stringently, we show the same wall persists when the ground truth is
generated by a genuinely non-bi-exponential velocity-dispersion mechanism, with an
off-model validity envelope characterizing how coverage degrades as the forward
model departs from the bi-exponential form. We additionally run the as-deployed
pipeline on real, publicly available multi-$b$ in-vivo DWI (ACRIN-6698 / I-SPY2
Breast DWI, The Cancer Imaging Archive, CC-BY-4.0): the deployment monitor correctly
fires on the synthetic-to-real transfer, and a label-free same-day test--retest
analysis ($n = 76$ tumors) shows the conformal interval width tracks scan--rescan
variability for the well-identified diffusion compartment. That section is
explicitly qualitative and makes no in-vivo coverage claim --- by design.

The work is built for reproducibility in the spirit of MRM Highlights: every numeric
claim traces to a gated, seeded checkpoint; conditional results carry empirical
confidence bands from a 16-seed re-run; and the complete pipeline, figures, and a
machine-checked consistency script are openly archived (Zenodo
10.5281/zenodo.20697755), regenerable from a single seed.

For full transparency, I disclose that I have a separate, related manuscript
currently under consideration at \emph{Magnetic Resonance in Medicine} [Manuscript
ID: MRM-26-27109], also concerning uncertainty quantification for IVIM parameter
mapping. The two are distinct in method and claim: the present work establishes
distribution-free conformal coverage and the pseudo-diffusion identifiability wall,
whereas the companion manuscript addresses the calibration of model-based IVIM
uncertainty. There is no overlapping content, and I am glad to provide the companion
manuscript to the editorial office on request.

This manuscript is original and has not been previously published. Other than the
related submission disclosed above, it is not under consideration elsewhere. I am
the sole author and declare no competing interests or specific funding.

Thank you for your consideration.

\medskip
\noindent Sincerely,

\medskip
\noindent Avery Karlin \\
The Annealing Signet Institute \\
Department of Applied Physics and Applied Mathematics, Columbia University, New
York, NY, USA \\
Department of Computer Science, University of Colorado Boulder, Boulder, CO, USA \\
ORCID: 0000-0003-3848-6782

\end{document}
