\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ebgaramond}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=teal,citecolor=teal,urlcolor=teal]{hyperref}

\input{numbers.tex}
\emergencystretch=2em   % avoid overfull boxes with ebgaramond justification

\title{\textbf{Sextant: boundary-railing of conventional NLLS IVIM fits as an
assumption-free diagnostic}\\[2pt]
\large A bound-pinned pseudo-diffusion failure mode, re-centred from a calibration
ruler and replicated on open human-abdominal diffusion MRI}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Reported uncertainty in intravoxel incoherent motion (IVIM) imaging has been
critiqued as \emph{overextended}: a coverage or calibration statement is only as
trustworthy as the noise and forward model that generated its reference truth. We
answer that critique not by defending the calibration argument but by promoting a
result that \emph{cannot} be overextended because it assumes nothing. In a
conventional bounded non-linear least-squares (NLLS) bi-exponential IVIM fit, the
pseudo-diffusion coefficient $D^{*}$ is so weakly identified that the optimiser
frequently terminates \emph{on} a parameter bound. On an open human-abdominal
acquisition, \HomoFrac\% of high-SNR voxels (n${=}$\HomoN) have a bound-railed NLLS
$D^{*}$ (95\% bootstrap CI [\HomoCIlo, \HomoCIhi]), reproducing a previously
reported figure; across the full abdominal ROI the rate is \FullFrac\%
([\FullCIlo, \FullCIhi], n${=}$\FullN). It survives deliberately generous bounds
(\HomoWide\% / \FullWide\%), persists across SNR, and replicates on an independent
liver-DWI cohort (TCGA-LIHC, \lihcLicense): \LihcFourFrac\% at the clean 4-b scheme
and \LihcThreeFrac\% across \LihcThreeSubj{} subjects at a sparse 3-b scheme.
Railing is a fact about the optimiser and the data --- no ground truth, no
noise-model trust --- and so is immune to the overextension objection. The
calibration ruler is retained but demoted to a scoped secondary diagnostic,
explicitly limited to data with ground truth.
\end{abstract}

\section{Introduction}
The clinically load-bearing question in quantitative diffusion MRI is not how
accurate a point estimate is, but whether a reported \emph{error bar} can be
believed. A natural way to answer it is a calibration ruler --- coverage, expected
calibration error, sharpness at fixed nominal levels --- that scores whether a
method's stated uncertainty is honest. The objection a sceptical reviewer raises
is that such a ruler is itself \emph{overextended}: empirical coverage is computed
against a reference truth simulated under an assumed noise and forward model, so a
coverage claim inherits exactly the trust it was supposed to certify.

This paper takes the objection seriously and re-aims the contribution. We promote
to the primary claim a phenomenon that requires no calibration trust at all:
\textbf{boundary-railing} of the conventional NLLS IVIM fit. Because railing is
read directly off the optimiser's output, it carries no calibration assumption and
therefore cannot be overextended. The calibration ruler is not discarded --- it is
kept as a clearly scoped \emph{secondary} diagnostic (Section~\ref{sec:ruler}).

\section{Boundary-railing}
The bi-exponential IVIM model is
\[
S(b)/S_0 = (1-f)\,e^{-bD} + f\,e^{-bD^{*}},
\]
fit per voxel by bounded NLLS over $(D, D^{*}, f)$. The pseudo-diffusion term
$D^{*}$ governs the fast perfusion decay seen only at low $b$; it is the
ill-conditioned parameter. When the likelihood is flat in $D^{*}$, a
trust-region optimiser does not find an interior optimum --- it slides to a box
constraint and stops. We call a voxel \emph{railed} if its NLLS $D^{*}$ lands
within $\varepsilon$ of either bound. The railed fraction among high-SNR voxels is
our statistic. It needs no ground truth: it is a property of the estimator applied
to the data.

\section{Methods}
\textbf{Reuse, not reimplementation.} The NLLS fit, the per-voxel signal/SNR
reduction, and the rail thresholds are reused read-only from the prior IVIM
calibration work (the \texttt{fashion\_reuse} module loads the exact functions by
source extraction). We add the bootstrap, the SNR/rail-direction characterisation,
the full-ROI generalisation, and the independent replication.

\textbf{Data.} The primary cohort is the OSIPI TF2.4 open human-abdominal IVIM
acquisition (Philips 3T; Zenodo 10.5281/zenodo.14605039, CC-BY-4.0), analysed at
its curated homogeneous ROI (the original cohort) and at the full abdominal ROI.
The independent cohort is TCGA-LIHC liver DWI (TCIA, \lihcLicense, DOI
\lihcDOI{}; \lihcNseries{} series), an unrelated site, scanner (Siemens 1.5T), and
organ. All imaging is download-on-demand; only provenance manifests are committed.

\textbf{Pre-registered thresholds.} Fixed before running: a cohort
\emph{replicates} if the railed fraction is $\ge\thrPoint\%$ \emph{and} its 95\%
bootstrap lower bound is $\ge\thrCIlo\%$; otherwise the claim scopes down. Bootstrap
is voxel-level resampling, \nboot{} resamples, seed \seed.

\section{Results}
\subsection{Primary: the original cohort}
On the curated homogeneous ROI, \HomoFrac\% of the \HomoN{} high-SNR voxels have a
bound-railed NLLS $D^{*}$ (95\% CI [\HomoCIlo, \HomoCIhi]), matching the previously
reported \fashionReported\% on the same data. Verdict: \HomoVerdict.

\subsection{Generalisation: the full abdomen}
Extending from the curated ROI to the entire abdominal ROI (n${=}$\FullN), the rate
is \FullFrac\% (95\% CI [\FullCIlo, \FullCIhi]); verdict \FullVerdict. Railing is
therefore not a property of a hand-chosen homogeneous patch.

\subsection{Independent replication: TCGA-LIHC liver DWI}
On an independent liver-DWI cohort the phenomenon recurs. At the clean 4-b scheme
($b{=}0,50,500,800$) the railed fraction is \LihcFourFrac\% (95\% CI [\LihcFourCIlo,
\LihcFourCIhi], n${=}$\LihcFourN; verdict \LihcFourVerdict). Across \LihcThreeSubj{}
subjects at a sparse 3-b clinical scheme ($b{=}50,400,800$, no $b{=}0$; normalised
by the lowest $b$, hence approximate) the pooled fraction is \LihcThreeFrac\%
([\LihcThreeCIlo, \LihcThreeCIhi], n${=}$\LihcThreeN; verdict \LihcThreeVerdict).
The sparser the perfusion sampling, the worse $D^{*}$ identifiability, and the more
the fit rails --- the expected direction.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{lrrrl}
\toprule
Cohort & $n$ (high-SNR) & Railed & 95\% CI & Verdict\\
\midrule
Abdomen, homogeneous ROI (OSIPI) & \HomoN & \HomoFrac\% & [\HomoCIlo, \HomoCIhi] & \HomoVerdict\\
Abdomen, full ROI (OSIPI) & \FullN & \FullFrac\% & [\FullCIlo, \FullCIhi] & \FullVerdict\\
Liver, 4-b $0/50/500/800$ (TCGA-LIHC) & \LihcFourN & \LihcFourFrac\% & [\LihcFourCIlo, \LihcFourCIhi] & \LihcFourVerdict\\
Liver, 3-b $50/400/800$ (TCGA-LIHC) & \LihcThreeN & \LihcThreeFrac\% & [\LihcThreeCIlo, \LihcThreeCIhi] & \LihcThreeVerdict\\
\bottomrule
\end{tabular}
\caption{Boundary-railing of NLLS $D^{*}$ across open human-abdominal cohorts
(seed \seed, \nboot-bootstrap). Pre-registered thresholds: railed
$\ge\thrPoint\%$ with CI lower bound $\ge\thrCIlo\%$.}
\end{table}

\subsection{When and why it rails}
Railing does not vanish with SNR --- it persists across every SNR stratum --- which
identifies it as \emph{structural} weak identifiability of $D^{*}$ rather than mere
measurement noise. It is dominated by the \emph{upper} bound (the homogeneous ROI
splits \HomoLower\% lower / \HomoUpper\% upper; the full ROI \FullLower\% /
\FullUpper\%): the high-$D^{*}$ identifiability wall, where the perfusion compartment
is so weakly constrained the optimiser is pushed to the ceiling. Under deliberately
generous bounds the rate falls but stays a substantial minority (\HomoWide\% /
\FullWide\%), so railing is not an artefact of tight bounds.

\subsection{Secondary: the scoped calibration ruler}
\label{sec:ruler}
The calibration ruler (coverage / ECE / sharpness) is retained as a secondary
diagnostic, reused read-only. Its scope is explicit: it scores estimates against a
known truth and is therefore \emph{undefined on the real abdominal scan}, which has
no ground-truth $(D, D^{*}, f)$. That limitation --- the very point the
overextension critique raised --- is the argument for leading with railing, which
needs no truth.

\section{Differentiation from learned uncertainty (Casali et al., 2026)}
Casali et al.\ (\emph{NMR in Biomedicine}, 2026; arXiv:2508.04588) propose a
\emph{method} --- deep-ensemble mixture-density networks for IVIM uncertainty ---
validated in vivo on \textbf{mouse brain}, and themselves report residual $D^{*}$
overconfidence. Sextant is not a competing UQ method; it documents an
assumption-free failure mode of the conventional fit on \textbf{human-abdominal}
tissue, where perfusion (and thus $D^{*}$) is clinically meaningful. The
differentiation is strictly empirical and scoped to the data: human-abdominal DWI
reaches a regime that preclinical mouse-brain validation does not, and Casali's own
$D^{*}$ overconfidence is the learned echo of the same weak-identifiability wall
that makes the conventional fit rail.

\section{Limitations and scope}
The open primary cohort is a single subject; we make no population claim from it,
reporting a within-subject voxel statistic with a voxel-bootstrap CI and supporting
generality by cross-cohort agreement (homogeneous, full, and independent liver). The
3-b liver scheme lacks $b{=}0$ and is normalised by its lowest $b$, an approximation
flagged as such; the 4-b liver cohort with $b{=}0$ is the clean independent test.
Railing magnitude depends on the bound width and the b-scheme density; we report
both the tight (original) and wide (generous) bounds and the per-scheme results
rather than a single number.

\section{Conclusion}
A large, reproducible fraction of conventional NLLS IVIM fits rail to a parameter
bound on open human-abdominal data, and the effect generalises across ROI, bounds,
SNR, and an independent liver cohort. Because it is read off the optimiser without
any calibration-trust assumption, it is the bulletproof primary claim the
overextension critique demanded; the calibration ruler remains useful but scoped.

\section*{Reproducibility}
One command, \texttt{bash Sextant/reproduce.sh}, fetches the open data
(MD5/checksum-verified; raw arrays never committed), runs the seeded analysis and
tests, and rebuilds this manuscript through the consistency gate that regenerates
every number above from \texttt{results/railing\_results.json}.

\end{document}
